---
title: "Regression Models - Project"
author: "justmzen"
date: "8/21/2020"
output: pdf_document
---

# Executive summary

The aim of the following study is to evaluate the relationship among many variables and miles per gallon (mpg) as the outcome. The main questions are:
- Is an automatic or manual ammission better for mtcars?
- Quantify the mtcars difference between automatic and manual ammissions.

# Data loading

```{r, echo = TRUE, message = FALSE}
library('datasets'); data(mtcars)
```

# Exploratory data analysis

The scope of the exploratory data analysis is to find any interesting relationship among the outcome and the possible regressors, including the ammission type. The detailed analysis is in the Appendix.

# Appendix

## Exploratory data analysis

```{r, echo = TRUE}

```

```{r, echo = TRUE}

```

```{r, echo = TRUE}

```

```{r, echo = TRUE}

```

```{r, echo = TRUE}

```

```{r, echo = TRUE}

```

```{r, echo = TRUE}

```

```{r, echo = TRUE}

```










## Transmission

```{r, echo = TRUE, fig.width=5, fig.height=3, fig.align='center'}
qplot(factor(am), mpg, data=mtcars, geom='boxplot', fill=factor(am))
```

There are many subcategories for the am variable. For this reason, a new categorical variable called "Transmission" will be included in the data set with two levels, automatic and manual:

```{r, echo = TRUE}
mtcars$ammission <- as.factor(ifelse(grepl(pattern = '^auto',
                                           x = mtcars$am) == TRUE,
                                     'automatic',
                                     'manual'))
```

## Mileage in city

```{r, echo = TRUE, fig.width=4, fig.height=3, fig.align='center'}
summary(mtcars$cty)
qplot(cty, data=mtcars, geom="histogram", bins=25)
```

# Comparing automatic and manual ammissions

```{r, echo = TRUE, fig.width=4, fig.height=3, fig.align='center'}
qplot(ammission, cty, data=mtcars, geom="boxplot", fill=ammission, ylab = 'City miles per gallon')
```

The plot suggests that manual ammission is better than automatic ammission based on the cty variable.

A t-test can verify the mean of manual ammission is greater than the mean of automatic ammission.

```{r, echo = TRUE}
t.test(mtcars[ammission == 'manual','cty'],
       mtcars[ammission == 'automatic','cty'],
       alternative = "greater", var.equal = FALSE)$p.value
```

Since the p-value is very low, it is possible to reject the null hypothesis that the means are equal and accept the alternative hypothesis, which is that the mean of the manual ammission is greater than the automatic one.

# Models

```{r, echo = TRUE}
fit1 <- lm(cty ~ ammission, data = mtcars)
summary(fit1)

fit2 <- update(fit1, cty ~ ammission + class -1)
summary(fit2)

fit3 <- update(fit1, cty ~ ammission + class + fl)
summary(fit3)

fit4 <- update(fit1, cty ~ ammission + class + fl + displ)
fit5 <- update(fit1, cty ~ ammission + class + fl + displ + drv)
fitAll <- update(fit1, cty ~ ammission + class + fl + displ + drv +
                         manufacturer + model + year)
```

# Model selection

```{r, echo = TRUE, message = FALSE}
anova(fit1, fit2, fit3, fit4, fit5, fitAll)

require(car)

```

# Interpretation of the coefficients

```{r, echo = TRUE}

```

# Diagnostics
Diagnostics and/or residual plots

```{r, echo = TRUE}
# e <- resid(fit)
# plot(fit)
```

# Conclusion
Inference / uncertainty quantification

```{r, echo = TRUE}
# Confidence intervals
# sumCoef <- summary(fit)$coefficients
# sumCoef[1,1] + c(-1,1)*qt(.975, df=fit$df)*sumCoef[1,2]
# sumCoef[2,1] + c(-1,1)*qt(.975, df=fit$df)*sumCoef[2,2]
```
